---
title: "Project_3"
author: "Soniya Shahi"
date: "2024-11-05"
output:
  pdf_document: default
  html_document: default
  word_document: default
---


```{r message=FALSE, warning=FALSE}
#the necessary libraries used in the project
library(readxl)
suppressMessages(library(dplyr))
suppressMessages(library(plotly))
library(ggplot2)
library(GGally)
```

## Data Preparation

```{r}
#converting the excel data to R dataset
unknown_data <- read_excel("Vole Skulls.xlsm", sheet ="Unknown")
subterraneus_data <- read_excel("Vole Skulls.xlsm", sheet ="Subterraneus")
multiplex_data <- read_excel("Vole Skulls.xlsm", sheet = "Multiplex")
```


```{r}
unknown_data<- unknown_data %>%
rename(
    "Index" = "Vole Skull Index" ,
    "ID" = "Chromosomal Analysis ID",
    "Length" = "Condylo Incisive Length or Skull Length (0.01mm)",
    "Height" = "Skull Height Above Bullae (0.01mm)",
    "Width" = "Skull Width Across Rostrum (0.01mm)"
  )

subterraneus_data<- subterraneus_data %>%
rename(
    "Index" = "Vole Skull Index" ,
    "ID" = "Chromosomal Analysis ID",
    "Length" = "Condylo Incisive Length or Skull Length (0.01mm)",
    "Height" = "Skull Height Above Bullae (0.01mm)",
    "Width" = "Skull Width Across Rostrum (0.01mm)"
  )

multiplex_data<- multiplex_data %>%
rename(
    "Index" = "Vole Skull Index" ,
    "ID" = "Chromosomal Analysis ID",
    "Length" = "Condylo Incisive Length or Skull Length (0.01mm)",
    "Height" = "Skull Height Above Bullae (0.01mm)",
    "Width" = "Skull Width Across Rostrum (0.01mm)"
  )
```


## Exploratory Data Analysis

```{r}
#number of rows remaining in unknown_data
raw_remaining_rows <- nrow(unknown_data)
print(raw_remaining_rows)

```

Missing values for Subterraneus data (Missing values showing in 12th row
even though correctly imported)

```{r}
missing_values <- sapply(subterraneus_data, function(x) sum(is.na(x)))
# Print the result
print(missing_values)
```

Missing values for Multiples data (Missing values showing in 33th row
even though correctly imported)

```{r}
missing_values <- sapply(multiplex_data, function(x) sum(is.na(x)))
# Print the result
print(missing_values)
```

```{r}
missing_values <- sapply(unknown_data, function(x) sum(is.na(x)))
# Print the result
print(missing_values)
```
```{r}
# Check for duplicates in the entire dataframe
duplicates <- duplicated(multiplex_data)

# Print rows that are duplicates
duplicate_rows <- multiplex_data[duplicates, ]
print(duplicate_rows)

```
```{r}
#  Delete duplicates in the entire dataframe
multiplex_data <- distinct(multiplex_data)

```


We then omitted the missing values from both the data frame.

```{r}
# Remove rows with missing values
subterraneus_data <- na.omit(subterraneus_data)
multiplex_data <- na.omit(multiplex_data)
```

We then checked outlier for the data frames. Starting with Subterraneus
Data.

```{r}
#merge data sets, analysis id, length, height and width

Orginal_merged_data <- multiplex_data %>% full_join(subterraneus_data, by = c("ID", "Length" , "Height", "Width" ))
ggpairs(Orginal_merged_data,columns = c("Length", "Height", "Width"),mapping = aes(color = ID, alpha = 0.2))
```


```{r message=FALSE, warning=FALSE}
# Create 3D scatter plot with specified columns
scatter3d <- plot_ly(subterraneus_data, x = ~subterraneus_data$Length, y = ~subterraneus_data$`Width`, z = ~subterraneus_data$Height, type = "scatter3d", mode = "markers")

# Customize layout directly in the plot_ly function
scatter3d <- scatter3d %>% layout(
  scene = list(
    xaxis = list(title = 'Skull Length'),
    yaxis = list(title = 'Skull Width'),
    zaxis = list(title = 'Skull Height')
  )
)

# Display the plot
scatter3d
```

Now handling the outliers --\> setting them to the 97.5th percentile
value. Upper bound set to 97.5th percentile Lower bound set to 2.5th
percentile

```{r}
# Calculate IQR-based lower and upper bounds for x, y, and z
Q1_x <- quantile(subterraneus_data[[3]], 0.25)
Q3_x <- quantile(subterraneus_data[[3]], 0.75)
IQR_x <- Q3_x - Q1_x

Q1_y <- quantile(subterraneus_data[[4]], 0.25)
Q3_y <- quantile(subterraneus_data[[4]], 0.75)
IQR_y <- Q3_y - Q1_y

Q1_z <- quantile(subterraneus_data[[5]], 0.25)
Q3_z <- quantile(subterraneus_data[[5]], 0.75)
IQR_z <- Q3_z - Q1_z

# Replace outliers with median for x
median_x <- median(subterraneus_data[[3]], na.rm = TRUE)
subterraneus_data[[3]][subterraneus_data[[3]] < (Q1_x - 1.5 * IQR_x)] <- median_x
subterraneus_data[[3]][subterraneus_data[[3]] > (Q3_x + 1.5 * IQR_x)] <- median_x

# Replace outliers with median for y
median_y <- median(subterraneus_data[[4]], na.rm = TRUE)
subterraneus_data[[4]][subterraneus_data[[4]] < (Q1_y - 1.5 * IQR_y)] <- median_y
subterraneus_data[[4]][subterraneus_data[[4]] > (Q3_y + 1.5 * IQR_y)] <- median_y

# Replace outliers with median for z
median_z <- median(subterraneus_data[[5]], na.rm = TRUE)
subterraneus_data[[5]][subterraneus_data[[5]] < (Q1_z - 1.5 * IQR_z)] <- median_z
subterraneus_data[[5]][subterraneus_data[[5]] > (Q3_z + 1.5 * IQR_z)] <- median_z

```

Now checking the data distribution after the outlier is manipulated and
replaced.

```{r message=FALSE, warning=FALSE}
# Create 3D scatter plot with specified columns
scatter3d <- plot_ly(subterraneus_data, x = ~subterraneus_data$Length, y = ~subterraneus_data$`Width`, z = ~subterraneus_data$Height, type = "scatter3d", mode = "markers")

# Customize layout directly in the plot_ly function
scatter3d <- scatter3d %>% layout(
  scene = list(
    xaxis = list(title = 'Skull Length'),
    yaxis = list(title = 'Skull Width'),
    zaxis = list(title = 'Skull Height')
  )
)

# Display the plot
scatter3d
```

Now we can clearly see that all the data poins fall under the same
cluster.

We now repeat the same process of finding the oultiers and replacing
them with upper bound value and lower bound value for the multiplex
dataframe.

```{r message=FALSE, warning=FALSE}

# Create 3D scatter plot with specified columns
scatter3d <- plot_ly(multiplex_data, x = ~multiplex_data[[3]], y = ~multiplex_data[[4]], z = ~multiplex_data[[5]], type = "scatter3d", mode = "markers")

# Customize layout directly in the plot_ly function
scatter3d <- scatter3d %>% layout(
  scene = list(
    xaxis = list(title = 'Skull Length'),
    yaxis = list(title = 'Skull Width'),
    zaxis = list(title = 'Skull Height')
  )
)
# Display the plot
scatter3d

```

```{r}
# Calculate the 2.5th and 97.5th percentiles for x, y, and z
Q1_x <- quantile(multiplex_data[[3]], 0.025)
Q3_x <- quantile(multiplex_data[[3]], 0.975)
IQR_x <- Q3_x - Q1_x

Q1_y <- quantile(multiplex_data[[4]], 0.025)
Q3_y <- quantile(multiplex_data[[4]], 0.975)
IQR_y <- Q3_y - Q1_y

Q1_z <- quantile(multiplex_data[[5]], 0.025)
Q3_z <- quantile(multiplex_data[[5]], 0.975)
IQR_z <- Q3_z - Q1_z

# Replace outliers with median for x
median_x <- median(multiplex_data[[3]], na.rm = TRUE)
multiplex_data[[3]][multiplex_data[[3]] < (Q1_x - 1.5 * IQR_x)] <- median_x
multiplex_data[[3]][multiplex_data[[3]] > (Q3_x + 1.5 * IQR_x)] <- median_x

# Replace outliers with median for y
median_y <- median(multiplex_data[[4]], na.rm = TRUE)
multiplex_data[[4]][multiplex_data[[4]] < (Q1_y - 1.5 * IQR_y)] <- median_y
multiplex_data[[4]][multiplex_data[[4]] > (Q3_y + 1.5 * IQR_y)] <- median_y

# Replace outliers with median for z
median_z <- median(multiplex_data[[5]], na.rm = TRUE)
multiplex_data[[5]][multiplex_data[[5]] < (Q1_z - 1.5 * IQR_z)] <- median_z
multiplex_data[[5]][multiplex_data[[5]] > (Q3_z + 1.5 * IQR_z)] <- median_z


# Replace outliers with median for z
multiplex_data[[5]][multiplex_data[[5]] < Q1_z] <- median_z
multiplex_data[[5]][multiplex_data[[5]] > Q3_z] <- median_z

```

```{r message=FALSE, warning=FALSE}

# Create 3D scatter plot with specified columns
scatter3d <- plot_ly(multiplex_data, x = ~multiplex_data[[3]], y = ~multiplex_data[[4]], z = ~multiplex_data[[5]], type = "scatter3d", mode = "markers")

# Customize layout directly in the plot_ly function
scatter3d <- scatter3d %>% layout(
  scene = list(
    xaxis = list(title = 'Skull Length'),
    yaxis = list(title = 'Skull Width'),
    zaxis = list(title = 'Skull Height')
  )
)
# Display the plot
scatter3d

```

Now we check the data distribution for both Subterraneus and Multiplex
data. First we merge them and then check the distribution.

```{r}
# we have two data set known chromosomal analysis and one with unknown
subterraneus_data <- subterraneus_data[,-1]
multiplex_data <-multiplex_data[,-1]
#merge data sets, analysis id, length, height and width 
merged_data <- multiplex_data %>% full_join(subterraneus_data, by = c("ID", "Length" , "Height", "Width" ))

```

```{r message=FALSE, warning=FALSE}

ggpairs(
  merged_data, 
  columns = c("Length", "Height", "Width"), # data columns
  mapping = aes(color = ID, alpha = 0.2) # color by chromosomal ID
)
```

Examining the data, we observe scattered values for each variable, with
intersections between certain values. Notably, there is overlap between classes.


```{r}
unknown_data<-unknown_data[complete.cases(unknown_data[ ,"Index"]), ]
unknown_data
```

We will be using  performing cross-validation for multiple linear regression models to assess their performance.

```{r}
# Set seed for reproducibility
set.seed(123)

# Filter rows without missing values
complete_data <- na.omit(unknown_data)

# Define the number of folds for cross-validation
num_folds <- 5

# Create a data frame to store cross-validation results
cv_results <- data.frame(Model = character(), R_squared = numeric(), MAPE = numeric(), stringsAsFactors = FALSE)

# List of models
models <- list(
  "Width from Length and Height" = "Width ~ Length + Height",
  "Width from Length" = "Width ~ Length",
  "Width from Height" = "Width ~ Height",
  "Height from Length and Width" = "Height ~ Length + Width",
  "Height from Length" = "Height ~ Length",
  "Height from Width" = "Height ~ Width"
)


# Perform k-fold cross-validation for each model
for (model_name in names(models)) {
  formula <- as.formula(models[[model_name]])
  
  # Create a model using lm
  model <- lm(formula, data = complete_data)
  
  # Access R-squared value from the summary of the linear model
  r_squared <- sprintf("%.3f", summary(model)$r.squared)

  # Make predictions on the entire dataset
  predictions <- predict(model, newdata = complete_data)
  
  # Extract response variable using model.frame
  response_variable <- all.vars(formula)[1]
  
  # Calculate MAPE with three decimal places
  mape <- sprintf("%.3f", mean(abs((predictions - complete_data[[response_variable]]) / complete_data[[response_variable]])) * 100)

  # Append results to the data frame
  cv_results <- rbind(cv_results, data.frame(Model = model_name, R_squared = as.numeric(r_squared), MAPE = as.numeric(mape)))
}

# Print the cross-validation results
print(cv_results)



```

Now we merge the unknown data with all the data.

```{r}
merged_data <- merged_data %>% full_join(unknown_data, by = c("ID", "Length" , "Height", "Width" ))
```

We repeat the similar process to check for the outlier and replace it
with upper bound and lower bound.

```{r message=FALSE, warning=FALSE}
# Create 3D scatter plot with specified columns
scatter3d <- plot_ly(merged_data, x = ~merged_data[[2]], y = ~merged_data[[3]], z = ~merged_data[[4]], type = "scatter3d", mode = "markers")

# Customize layout directly in the plot_ly function
scatter3d <- scatter3d %>% layout(
  scene = list(
    xaxis = list(title = 'Skull Length'),
    yaxis = list(title = 'Skull Width'),
    zaxis = list(title = 'Skull Height')
  )
)
# Display the plot
scatter3d
```

```{r}
# Calculate the 2.5th and 97.5th percentiles for x, y, and z
Q1_x <- quantile(merged_data[[2]], 0.025, na.rm = TRUE)
Q3_x <- quantile(merged_data[[2]], 0.975, na.rm = TRUE)
IQR_x <- Q3_x - Q1_x

Q1_y <- quantile(merged_data[[3]], 0.025, na.rm = TRUE)
Q3_y <- quantile(merged_data[[3]], 0.975, na.rm = TRUE)
IQR_y <- Q3_y - Q1_y

Q1_z <- quantile(merged_data[[4]], 0.025, na.rm = TRUE)
Q3_z <- quantile(merged_data[[4]], 0.975, na.rm = TRUE)
IQR_z <- Q3_z - Q1_z

# Replace outliers with median for x
median_x <- median(merged_data[[2]], na.rm = TRUE)
merged_data[[2]][merged_data[[2]] < (Q1_x - 1.5 * IQR_x)] <- median_x
merged_data[[2]][merged_data[[2]] > (Q3_x + 1.5 * IQR_x)] <- median_x

# Replace outliers with median for y
median_y <- median(merged_data[[3]], na.rm = TRUE)
merged_data[[3]][merged_data[[3]] < (Q1_y - 1.5 * IQR_y)] <- median_y
merged_data[[3]][merged_data[[3]] > (Q3_y + 1.5 * IQR_y)] <- median_y

# Replace outliers with median for z
median_z <- median(merged_data[[4]], na.rm = TRUE)
merged_data[[4]][merged_data[[4]] < (Q1_z - 1.5 * IQR_z)] <- median_z
merged_data[[4]][merged_data[[4]] > (Q3_z + 1.5 * IQR_z)] <- median_z
```

```{r message=FALSE, warning=FALSE}
# Create 3D scatter plot with specified columns
scatter3d <- plot_ly(merged_data, x = ~merged_data[[2]], y = ~merged_data[[3]], z = ~merged_data[[4]], type = "scatter3d", mode = "markers")

# Customize layout directly in the plot_ly function
scatter3d <- scatter3d %>% layout(
  scene = list(
    xaxis = list(title = 'Skull Length'),
    yaxis = list(title = 'Skull Width'),
    zaxis = list(title = 'Skull Height')
  )
)
# Display the plot
scatter3d
```

```{r message=FALSE, warning=FALSE}
ggpairs(merged_data, ggplot2::aes(colour=`ID`))
```

Now, from the above graph we can clearly see the unknown data has a
significant cluster in between the overlapping intersection between
multiplex and subterraneus data. Hence, that niche area will be very
difficult to predict.

Now we use cross validation with 5 fold as a function to train and test
the model.

```{r}

perform_cross_validation <- function(data, formula) {
  # Convert 'ID' to a factor
  data$ID <- as.factor(data$ID)

  # Filter data for training and cross-validation (excluding 'unknown')
  filtered_data <- data[data$ID != "unknown", ]

  # Set a seed for reproducibility
  set.seed(123)

  # Split data into 5 folds
  folds <- cut(seq(1, nrow(filtered_data)), breaks = 5, labels = FALSE)

  # Initialize vectors to store results
  train_accuracies <- numeric(length = 5)
  valid_accuracies <- numeric(length = 5)

  # Perform 5-fold cross-validation
  for (i in 1:5) {
    # Create indices for training and validation sets
    train_indices <- which(folds != i)
    valid_indices <- which(folds == i)

    # Fit the model on the training set
    model <- glm(formula, data = filtered_data[train_indices, ], family = "binomial")

    # Make predictions on the training set
    train_predictions <- predict(model, newdata = filtered_data[train_indices, ], type = "response") > 0.5
    train_predictions <- ifelse(train_predictions, "subterraneus", "multiplex")

    # Make predictions on the validation set
    valid_predictions <- predict(model, newdata = filtered_data[valid_indices, ], type = "response") > 0.5
    valid_predictions <- ifelse(valid_predictions, "subterraneus", "multiplex")

    # Calculate and store accuracies
    train_accuracies[i] <- mean(train_predictions == filtered_data$ID[train_indices])
    valid_accuracies[i] <- mean(valid_predictions == filtered_data$ID[valid_indices])
  }

  # Calculate mean accuracies
  mean_train_accuracy <- mean(train_accuracies)
  mean_valid_accuracy <- mean(valid_accuracies)

  # Return the results
  return(list(
    train_accuracies = train_accuracies,
    valid_accuracies = valid_accuracies,
    mean_train_accuracy = mean_train_accuracy,
    mean_valid_accuracy = mean_valid_accuracy
  ))
}
```

The following chuck will takes in merged data train and test the 7
model
Length + Width + Height Length + Width Length + Height Width + Height
Length Width Height

```{r}

# Convert 'ID' to a factor
merged_data$ID <- as.factor(merged_data$ID)

# Define a vector of formulas
formulas <- c(
  "ID ~ Length + Width + Height",
  "ID ~ Length + Width",
  "ID ~ Length + Height",
  "ID ~ Width + Height",
  "ID ~ Length",
  "ID ~ Width",
  "ID ~ Height"
)

# Initialize a dataframe to store the results
results_df <- data.frame(Formula = character(), Mean_Training_Accuracy = numeric(), Mean_Validation_Accuracy = numeric(), stringsAsFactors = FALSE)

# Perform cross-validation for each formula
for (formula_text in formulas) {
  formula <- as.formula(formula_text)
  
  # Perform cross-validation using the function
  results <- perform_cross_validation(merged_data, formula)
  
  # Add results to the dataframe
  results_df <- rbind(results_df, data.frame(Formula = formula_text, Mean_Training_Accuracy = results$mean_train_accuracy, Mean_Validation_Accuracy = results$mean_valid_accuracy))
}

# Print the results dataframe
print(results_df)
```

The model featuring only Width and Height exhibits the highest training
and validation accuracy relative to others. Its interpretation is more
straightforward compared to models with additional features. This
observation suggests that the inclusion of more features does not
necessarily guarantee an improvement in model accuracy.

***Stepwise Regression Model***

```{r}
# Assuming 'merged_data' is our merged dataset
library(boot)

# Convert 'ID' to a factor
merged_data$ID <- as.factor(merged_data$ID)

# Filter data for training (excluding 'unknown')
train_data <- merged_data[merged_data$ID != "unknown", ]

# Specify the formula for the logistic regression model
formula <- ID ~ Length + Height + Width + Length * Width + Length * Height + Width * Height

# Fit the stepwise logistic regression model using only training data
stepwise_model <- step(glm(formula, data = train_data, family = "binomial"), direction = "both")

# Display the final summary for the stepwise model
summary(stepwise_model)
```

The provided output illustrates a stepwise logistic regression process
with the goal of minimizing the Akaike Information Criterion (AIC) to
find the most parsimonious model for predicting the variable "ID." The
stepwise selection starts with a model including interactions between
Length, Height, and Width. At each step, the algorithm evaluates the
impact of removing or adding terms to the model based on AIC. The final
selected model is a logistic regression model with the formula "ID \~
Height + Width." The estimated coefficients indicate that both Height
and Width are statistically significant predictors of the response
variable "ID," with negative coefficients suggesting a negative impact
on the log-odds of the outcome. The resulting model has a residual
deviance of 48.086 on 87 degrees of freedom, and its AIC is 54.086. This
stepwise process highlights a streamlined model, emphasizing the
significance of Height and Width in predicting the binary outcome "ID."

```{r message=FALSE, warning=FALSE}
# Convert 'ID' to a factor
merged_data$ID <- as.factor(merged_data$ID)

# Coefficients of the line (replace with your actual coefficients)
intercept <- -77.71775/(-0.04253)
slope <- -(-0.09987) / (-0.04253)

# Create a scatterplot using ggplot2
ggplot(train_data, aes(x = Width, y = Height, color = ID)) +
  geom_point() +
  geom_abline(intercept = intercept, slope = slope, color = "black", linetype = "dashed", size = 1) +
  labs(title = "Scatterplot with Line",
       x = "Width",
       y = "Height") +
  scale_color_manual(values = c("subterraneus" = "blue", "multiplex" = "red")) +
  theme_minimal() +
  coord_cartesian(xlim = c(min(train_data$Width), max(train_data$Width)),
                  ylim = c(min(train_data$Height), max(train_data$Height)))
```

The provided figure illustrates a decision boundary, where data points
located above the line are classified as multiples, and those below are
classified as subterraneus. Additionally, it visually conveys that
predictions are not consistently accurate, as some points deviate from
their actual classifications. This is for the traning dataset.

```{r}
# Filter out rows where ID is not "unknown"
training_data <- merged_data[merged_data$ID != "unknown", ]

# Convert 'ID' to a factor
training_data$ID <- as.factor(training_data$ID)

# Initialize a list to store the models
models <- list()

# Train models for each formula
for (formula_text in formulas) {
  formula <- as.formula(formula_text)
  
  # Train the model using the training_data
  model <- glm(formula, data = training_data,family = binomial())
  
  # Store the model in the list
  models[[formula_text]] <- model
}

 # Function to convert probabilities to class labels
  convert_to_class <- function(probabilities, threshold = 0.5) {
    return(ifelse(probabilities > threshold, "subterraneus","multiplex"))
  }

```

```{r}
predict_from_datasets <- function(dataset, models) {
  # Omit datasets with missing values in all three columns "Length", "Width", and "Height"
 dataset <- dataset[!(is.na(dataset$Length) & is.na(dataset$Width) & is.na(dataset$Height)), ]
 
  # Split the dataset based on missing values
  dataset1 <- dataset[(is.na(dataset$Length) & !(is.na(dataset$Width)) & !(is.na(dataset$Height))) | complete.cases(dataset[c("Length", "Width", "Height")]), ]
  dataset2 <- dataset[is.na(dataset$Width) & complete.cases(dataset[c("Length", "Height")]), ]
  dataset3 <- dataset[is.na(dataset$Height) & complete.cases(dataset[c("Length", "Width")]), ]
  dataset4 <- dataset[is.na(dataset$Length) & is.na(dataset$Height) & !(is.na(dataset$Width)), ]
  dataset5 <- dataset[is.na(dataset$Height) & is.na(dataset$Width) & !(is.na(dataset$Length)), ]
  dataset6 <- dataset[is.na(dataset$Length) & is.na(dataset$Width) & !(is.na(dataset$Height)), ]
  

  # Make predictions and store in respective dataset's predicted_ID column
  dataset1$predicted_ID <- convert_to_class(predict(models[["ID ~ Width + Height"]], newdata = dataset1, type = "response"))
  dataset2$predicted_ID <- convert_to_class(predict(models[["ID ~ Length + Height"]], newdata = dataset2, type = "response"))
  dataset3$predicted_ID <- convert_to_class(predict(models[["ID ~ Length + Width"]], newdata = dataset3, type = "response"))
  dataset4$predicted_ID <- convert_to_class(predict(models[["ID ~ Width"]], newdata = dataset4, type = "response"))
  dataset5$predicted_ID <- convert_to_class(predict(models[["ID ~ Length"]], newdata = dataset5, type = "response"))
  dataset6$predicted_ID <- convert_to_class(predict(models[["ID ~ Height"]], newdata = dataset6, type = "response"))
  
 # Print the length of each dataset and the corresponding model used
  print(paste("No. of Data:", nrow(dataset1), ", Model Used: ID ~ Width + Height"))
  print(paste("   Subterraneus Count:", sum(dataset1$predicted_ID == "subterraneus")))
  print(paste("   Microtus Count:", sum(dataset1$predicted_ID == "multiplex")))
  
  print(paste("No. of Data:", nrow(dataset2), ", Model Used: ID ~ Length + Height"))
  print(paste("   Subterraneus Count:", sum(dataset2$predicted_ID == "subterraneus")))
  print(paste("   Microtus Count:", sum(dataset2$predicted_ID == "multiplex")))
  
  print(paste("No. of Data:", nrow(dataset3), ", Model Used: ID ~ Length + Width"))
  print(paste("   Subterraneus Count:", sum(dataset3$predicted_ID == "subterraneus")))
  print(paste("   Microtus Count:", sum(dataset3$predicted_ID == "multiplex")))
  
  print(paste("No. of Data:", nrow(dataset4), ", Model Used: ID ~ Width"))
  print(paste("   Subterraneus Count:", sum(dataset4$predicted_ID == "subterraneus")))
  print(paste("   Microtus Count:", sum(dataset4$predicted_ID == "multiplex")))
  
  print(paste("No. of Data:", nrow(dataset5), ", Model Used: ID ~ Length"))
  print(paste("   Subterraneus Count:", sum(dataset5$predicted_ID == "subterraneus")))
  print(paste("   Microtus Count:", sum(dataset5$predicted_ID == "multiplex")))
  
  print(paste("No. of Data:", nrow(dataset6), ", Model Used: ID ~ Height"))
  print(paste("   Subterraneus Count:", sum(dataset6$predicted_ID == "subterraneus")))
  print(paste("   Microtus Count:", sum(dataset6$predicted_ID == "multiplex")))
  # Merge all datasets and sort by Index
  merged_all <- rbind(dataset1, dataset2, dataset3, dataset4, dataset5, dataset6)
  merged_all <- merged_all[order(merged_all$Index), ]
  
  
  # Return the merged dataset
  return(merged_all)
}


```

```{r}
unknown_data <- predict_from_datasets(unknown_data, models)
unknown_data
```

```{r}
# Assuming 'merged_dataset' is your dataset
predicted_id_counts <- table(unknown_data$predicted_ID)

# Print the counts
print("Predicted_ID Counts:")
print(predicted_id_counts)

```