---
title: "Project_2"
author: "SAI RAM VODNALA"
date: "2024-11-09"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation

**EXCEL**

Removed the empty rows (35,14,22) and duplicate(29 multiplex) in data for multiplex, subterraneus and unknown sheets in excel. And changed column names then converted to CSV format.

## Exploratory Data Analysis

**R**

```{r}
#Importing the data sets
multiplex <- read.csv("Multiplex1.csv")
subterraneus <- read.csv("Subterraneus1.csv")
unknown <- read.csv("Unknown1.csv")
```

```{r}
#checking the top rows
head(multiplex)
head(subterraneus)
head(unknown)
```

```{r}
# Checking missing values
sum(is.na(multiplex))
sum(is.na(subterraneus))
```

```{r}
#checking the structure and data types
str(multiplex)
str(subterraneus)
str(unknown)
```
```{r}
summary(multiplex)
summary(subterraneus)
```

```{r}
#creating boxplots to find ouliers for Multiplex
library(ggplot2)

# Prepare data by melting it into long format for easy faceting
multiplex_long <- reshape2::melt(multiplex, id.vars = "vole_skull_index", measure.vars = c("skull_length", "skull_height", "skull_width"))

# Create box plots for all features
ggplot(multiplex_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +  # Red color for outliers
  scale_fill_brewer(palette = "Pastel1") +  # Colorful boxes using a pastel palette
  facet_wrap(~ variable, scales = "free_y") +  # Separate plots for each feature
  labs(title = "Box Plots of Multiplex Skull Measurements", x = "Measurement Type", y = "Measurement Value (0.01mm)") +
  theme_minimal() +
  theme(legend.position = "none",  # Hide legend since color distinguishes the features
        axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels for clarity

```

```{r}
#Creating Density charts to know distribution of a column values for Multiplex
library(ggplot2)

# Density plot for 'skull_length' in the multiplex dataset
p1_multiplex <- ggplot(multiplex, aes(x = skull_length)) + 
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density of Skull Length - Multiplex", x = "Skull Length", y = "Density") +
  theme_minimal()

# Density plot for 'skull_height' in the multiplex dataset
p2_multiplex <- ggplot(multiplex, aes(x = skull_height)) + 
  geom_density(fill = "red", alpha = 0.5) +
  labs(title = "Density of Skull Height - Multiplex", x = "Skull Height", y = "Density") +
  theme_minimal()

# Density plot for 'skull_width' in the multiplex dataset
p3_multiplex <- ggplot(multiplex, aes(x = skull_width)) + 
  geom_density(fill = "green", alpha = 0.5) +
  labs(title = "Density of Skull Width - Multiplex", x = "Skull Width", y = "Density") +
  theme_minimal()

# Display the plots
gridExtra::grid.arrange(p1_multiplex, p2_multiplex, p3_multiplex, ncol = 3)



```
The box plot and density plot show how the skull measurements—length, height, and width—vary for a species known as "Multiplex." Both skull length and height have a moderate spread with some rare, extreme values. In particular, skull width shows more variability, with values that frequently stray far from the average, both lower and higher. The density plots support this, showing that most skulls have lengths and heights close to a common middle value, but skull width varies more widely across different skulls. This broader variation in width is reflected in a spread-out pattern in its density plot, suggesting that skull width is a more diverse trait within this species.

```{r}
library(ggplot2)
#creating boxplots to find ouliers for subterraneus

# Prepare data by melting it into long format for easy faceting
subterraneus_long <- reshape2::melt(subterraneus, id.vars = "vole_skull_index", measure.vars = c("skull_length", "skull_height", "skull_width"))

# Create box plots for all features
ggplot(subterraneus_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +  # Red color for outliers
  scale_fill_brewer(palette = "Pastel1") +  # Colorful boxes using a pastel palette
  facet_wrap(~ variable, scales = "free_y") +  # Separate plots for each feature
  labs(title = "Box Plots of subterraneus Skull Measurements", x = "Measurement Type", y = "Measurement Value (0.01mm)") +
  theme_minimal() +
  theme(legend.position = "none",  # Hide legend since color distinguishes the features
        axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels for clarity




```

```{r}
#Creating Density charts to know distribution of a column values for subterraneus
library(ggplot2)

# Density plot for 'skull_length' in the subterraneus dataset
p1_subterraneus <- ggplot(subterraneus, aes(x = skull_length)) + 
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density of Skull Length - subterraneus", x = "Skull Length", y = "Density") +
  theme_minimal()

# Density plot for 'skull_height' in the subterraneus dataset
p2_subterraneus <- ggplot(subterraneus, aes(x = skull_height)) + 
  geom_density(fill = "red", alpha = 0.5) +
  labs(title = "Density of Skull Height - subterraneus", x = "Skull Height", y = "Density") +
  theme_minimal()

# Density plot for 'skull_width' in the subterraneus dataset
p3_subterraneus <- ggplot(subterraneus, aes(x = skull_width)) + 
  geom_density(fill = "green", alpha = 0.5) +
  labs(title = "Density of Skull Width - subterraneus", x = "Skull Width", y = "Density") +
  theme_minimal()

# Display the plots
gridExtra::grid.arrange(p1_subterraneus, p2_subterraneus, p3_subterraneus, ncol = 3)

```
box plots and density plots for Subterraneus skull measurements reveal significant outliers and extreme variability, particularly in skull length and height. The box plots show that while the median values for skull height and width are within a more typical range, the lengths and heights have extreme outliers far beyond the norm, with skull length outliers over 20,000 and skull height exceeding 8,000. The density plots confirm this distribution, displaying sharp peaks at lower values for height and width, indicating a concentration of most data at these levels, but with skull length showing a drastic spread, indicative of possible measurement errors or extraordinary specimens.

so we need to handle the outliers –> setting them to the 97.5th percentile value. Upper bound set to 97.5th percentile Lower bound set to 2.5th percentile
```{r}
# Calculate lower and upper bounds for x, y, and z
lower_bound_x <- quantile(multiplex[[3]], 0.025)
upper_bound_x <- quantile(multiplex[[3]], 0.975)

lower_bound_y <- quantile(multiplex[[4]], 0.025)
upper_bound_y <- quantile(multiplex[[4]], 0.975)

lower_bound_z <- quantile(multiplex[[5]], 0.025)
upper_bound_z <- quantile(multiplex[[5]], 0.975)

# Replace values outside the bounds for x, y, and z
multiplex[[3]][multiplex[[3]] < lower_bound_x] <- lower_bound_x
multiplex[[3]][multiplex[[3]] > upper_bound_x] <- upper_bound_x

multiplex[[4]][multiplex[[4]] < lower_bound_y] <- lower_bound_y
multiplex[[4]][multiplex[[4]] > upper_bound_y] <- upper_bound_y

multiplex[[5]][multiplex[[5]] < lower_bound_z] <- lower_bound_z
multiplex[[5]][multiplex[[5]] > upper_bound_z] <- upper_bound_z


# Calculate lower and upper bounds for x, y, and z
lower_bound_x <- quantile(subterraneus[[3]], 0.025)
upper_bound_x <- quantile(subterraneus[[3]], 0.975)

lower_bound_y <- quantile(subterraneus[[4]], 0.025)
upper_bound_y <- quantile(subterraneus[[4]], 0.975)

lower_bound_z <- quantile(subterraneus[[5]], 0.025)
upper_bound_z <- quantile(subterraneus[[5]], 0.975)

# Replace values outside the bounds for x, y, and z
subterraneus[[3]][subterraneus[[3]] < lower_bound_x] <- lower_bound_x
subterraneus[[3]][subterraneus[[3]] > upper_bound_x] <- upper_bound_x

subterraneus[[4]][subterraneus[[4]] < lower_bound_y] <- lower_bound_y
subterraneus[[4]][subterraneus[[4]] > upper_bound_y] <- upper_bound_y

subterraneus[[5]][subterraneus[[5]] < lower_bound_z] <- lower_bound_z
subterraneus[[5]][subterraneus[[5]] > upper_bound_z] <- upper_bound_z


```

```{r}
# Check the structure and summary of the cleaned data
multiplex_clean <- multiplex
subterraneus_clean <- subterraneus
summary(multiplex_clean)
summary(subterraneus_clean)


```

```{r}
# Check the structure and summary of the cleaned data
summary(multiplex_clean)
summary(subterraneus_clean)

```

```{r}
library(ggplot2)

# Prepare data by melting it into long format for easy faceting
multiplex_clean_long <- reshape2::melt(multiplex_clean, id.vars = "vole_skull_index", measure.vars = c("skull_length", "skull_height", "skull_width"))

# Create box plots for all features
ggplot(multiplex_clean_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +  # Red color for outliers
  scale_fill_brewer(palette = "Pastel1") +  # Colorful boxes using a pastel palette
  facet_wrap(~ variable, scales = "free_y") +  # Separate plots for each feature
  labs(title = "Box Plots of Multiplex_Clean Skull Measurements", x = "Measurement Type", y = "Measurement Value (0.01mm)") +
  theme_minimal() +
  theme(legend.position = "none",  # Hide legend since color distinguishes the features
        axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels for clarity
```

```{r}
library(ggplot2)

# Density plot for 'skull_length' in the multiplex_clean dataset
p1_multiplex_clean <- ggplot(multiplex_clean, aes(x = skull_length)) + 
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density of Skull Length - multiplex_clean", x = "Skull Length", y = "Density") +
  theme_minimal()

# Density plot for 'skull_height' in the multiplex_clean dataset
p2_multiplex_clean <- ggplot(multiplex_clean, aes(x = skull_height)) + 
  geom_density(fill = "red", alpha = 0.5) +
  labs(title = "Density of Skull Height - multiplex_clean", x = "Skull Height", y = "Density") +
  theme_minimal()

# Density plot for 'skull_width' in the multiplex_clean dataset
p3_multiplex_clean <- ggplot(multiplex_clean, aes(x = skull_width)) + 
  geom_density(fill = "green", alpha = 0.5) +
  labs(title = "Density of Skull Width - multiplex_clean", x = "Skull Width", y = "Density") +
  theme_minimal()

# Display the plots
gridExtra::grid.arrange(p1_multiplex_clean, p2_multiplex_clean, p3_multiplex_clean, ncol = 3)



```

```{r}
library(ggplot2)

# Prepare data by melting it into long format for easy faceting
subterraneus_clean_long <- reshape2::melt(subterraneus_clean, id.vars = "vole_skull_index", measure.vars = c("skull_length", "skull_height", "skull_width"))

# Create box plots for all features
ggplot(subterraneus_clean_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +  # Red color for outliers
  scale_fill_brewer(palette = "Pastel1") +  # Colorful boxes using a pastel palette
  facet_wrap(~ variable, scales = "free_y") +  # Separate plots for each feature
  labs(title = "Box Plots of subterraneus_clean Skull Measurements", x = "Measurement Type", y = "Measurement Value (0.01mm)") +
  theme_minimal() +
  theme(legend.position = "none",  # Hide legend since color distinguishes the features
        axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels for clarity

```

```{r}
library(ggplot2)

# Density plot for 'skull_length' in the subterraneus_clean dataset
p1_subterraneus_clean <- ggplot(subterraneus_clean, aes(x = skull_length)) + 
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density of Skull Length - subterraneus_clean", x = "Skull Length", y = "Density") +
  theme_minimal()

# Density plot for 'skull_height' in the subterraneus_clean dataset
p2_subterraneus_clean <- ggplot(subterraneus_clean, aes(x = skull_height)) + 
  geom_density(fill = "red", alpha = 0.5) +
  labs(title = "Density of Skull Height - subterraneus_clean", x = "Skull Height", y = "Density") +
  theme_minimal()

# Density plot for 'skull_width' in the subterraneus_clean dataset
p3_subterraneus_clean <- ggplot(subterraneus_clean, aes(x = skull_width)) + 
  geom_density(fill = "green", alpha = 0.5) +
  labs(title = "Density of Skull Width - subterraneus_clean", x = "Skull Width", y = "Density") +
  theme_minimal()

# Display the plots
gridExtra::grid.arrange(p1_subterraneus_clean, p2_subterraneus_clean, p3_subterraneus_clean, ncol = 3)



```

```{r}
# Specify the file path and save the multiplex_clean dataset as a CSV file
write.csv(multiplex_clean, "multiplex_clean.csv", row.names = FALSE)

# Specify the file path and save the subterraneus_clean dataset as a CSV file
write.csv(subterraneus_clean, "subterraneus_clean.csv", row.names = FALSE)

```

```{r}
library(dplyr)
# Merge the datasets
combined_data <- bind_rows(multiplex_clean, subterraneus_clean)
combined_data <- combined_data[, -1]
combined_data <- combined_data[, c("skull_length", "skull_height", "skull_width", "chromosomal_analysis_ID")]


# Save the combined dataset to a CSV file
write.csv(combined_data, "combined_data.csv", row.names = FALSE)
```

```{r}
library(dplyr)      # For data manipulation
library(ggplot2)    # For creating visualizations
library(gridExtra)  # For arranging multiple plots

```

```{r}
summary(combined_data)
```
## Training the data using CV
Now we use cross validation with 10 fold as a function to train and test the model.
```{R}
#creating seven feature combinations to develop glm models
library(caret)
combined_data$chromosomal_analysis_ID <- as.factor(combined_data$chromosomal_analysis_ID)

# Define feature combinations
feature_combinations <- list(
  c("skull_length", "skull_height", "skull_width"), # Model with three features
  c("skull_length", "skull_height"),                # Model with Length and Height
  c("skull_length", "skull_width"),                 # Model with Length and Rostrum
  c("skull_height", "skull_width"),                 # Model with Height and Rostrum
  c("skull_length"),                                # Model with Length only
  c("skull_height"),                                # Model with Height only
  c("skull_width")                                  # Model with Rostrum only
)

```

```{R}
#Training seven models
cv_results <- list()

for (i in seq_along(feature_combinations)) {
  # Define the formula based on feature combinations
  formula <- as.formula(paste("chromosomal_analysis_ID ~", paste(feature_combinations[[i]], collapse = "+")))
  
  # Set up cross-validation
  train_control <- trainControl(method = "cv", number = 10, classProbs = TRUE, savePredictions = "final")
  
  # Train the model using cross-validation
  cv_model <- train(formula, data = combined_data, method = "glm", family = binomial(), trControl = train_control)
  
  # Store results
  cv_results[[i]] <- cv_model
}

```

```{R}
# Print results from seven models
for (i in seq_along(cv_results)) {
  cat("\nResults for Model", i, "using Cross-Validation:\n")
  
  print(cv_results[[i]])
  print(summary(cv_results[[i]]))
}
```
##Predicting the unknown data
Making unknown data into seven subsets files based on features to test with their respective model.

```{r}
# Define all possible columns in the dataset for reference
all_columns <- c("skull_length", "skull_height", "skull_width")

# Define the feature combinations
feature_combinations <- list(
  c("skull_length", "skull_height", "skull_width"), # Model with three features
  c("skull_length", "skull_height"),                # Model with Length and Height
  c("skull_length", "skull_width"),                 # Model with Length and Rostrum
  c("skull_height", "skull_width"),                 # Model with Height and Rostrum
  c("skull_length"),                                # Model with Length only
  c("skull_height"),                                # Model with Height only
  c("skull_width")                                  # Model with Rostrum only
)

# Process each feature combination
for (i in seq_along(feature_combinations)) {
  # Select only the relevant columns for this model
  data_subset <- unknown[, feature_combinations[[i]], drop = FALSE]

  # Identify non-relevant columns for this combination
  non_relevant_columns <- setdiff(all_columns, feature_combinations[[i]])

  # Filter out rows where any of the non-relevant columns are not NA
  if (length(non_relevant_columns) > 0) {
    # Ensure that the subset operation does not drop dimensions, even if it's a single column
    temp_non_rel_data <- unknown[, non_relevant_columns, drop = FALSE]
    data_subset <- data_subset[rowSums(!is.na(temp_non_rel_data)) == 0, , drop = FALSE]
  }

  # Remove rows with any NAs in the relevant columns, ensuring data_subset stays as a dataframe
  if(ncol(data_subset) > 0) {
    data_subset <- data_subset[complete.cases(data_subset), , drop = FALSE]
  }

  # Check if the dataframe is empty or has been reduced incorrectly
  if (ncol(data_subset) > 0 && nrow(data_subset) > 0) {
    # Save to CSV only if data subset is not empty
    filename <- paste0("unknown_subset_", i, ".csv")
    write.csv(data_subset, filename, row.names = FALSE)
  } else {
    cat(sprintf("No data available to save for model %d due to filtering or NA removal.\n", i))
  }
}

```

predicting seven unknown data subsets with respective feature combination models.
```{R}
library(caret)  # If not already loaded

unknown_results <- list()

for (i in seq_along(cv_results)) {
  # Load the data for this model
  filename <- paste0("unknown_subset_", i, ".csv")
  model_data <- read.csv(filename)
  
  # Predict using the ith model
  predictions <- predict(cv_results[[i]], model_data, type = "raw")
  probabilities <- predict(cv_results[[i]], model_data, type = "prob")
  
  # Store results
  unknown_results[[i]] <- list(predictions = predictions, probabilities = probabilities)
}
```


Printing the predicted results
```{r}
# Summarize predictions
for (i in seq_along(unknown_results)) {
  cat("\nModel", i, "Summary:\n")
  print(table(unknown_results[[i]]$predictions))  # Show frequency of each predicted class
  
  if ("probabilities" %in% names(unknown_results[[i]])) {
    cat("\nModel", i, "Probability Summary:\n")
    summary_stats <- summary(unknown_results[[i]]$probabilities)  # Summarize probabilities if available
    print(summary_stats)
  }
}


```
