---
title: "Project_4"
author: "Sai Ram Vodnala, Bibek Adhikari"
date: "2025-05-04"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Installing required packages and loading libraries
```{r}
# Install packages, if not already installed
packages <- c("tidyverse", "randomForest", "cluster", "mclust", "ggplot2", "factoextra", "caret")
installed <- rownames(installed.packages())
to_install <- setdiff(packages, installed)
if(length(to_install)) install.packages(to_install)

# Load required libraries
library(tidyverse)
library(randomForest)
library(cluster)
library(mclust)
library(ggplot2)
library(factoextra)
library(caret)
library(readr)
```

#Loading the datasets
```{r}
# Loading unexploded, background and query data sets required for Project
setwd("D:/Spring-2025/Stat602/Project04")
uw <- read_csv("unexploded_watchlist.csv")
bg <- read.csv("background_data.csv")
q1 <- read.csv("query_1.csv")
q2 <- read.csv("query_2.csv")
q6 <- read.csv("query_6.csv")
```


# Previewing each of the loaded datasets
```{r}
#unexploded
head(uw)
```

```{r}
#background
head(bg)
```

```{r}
#query1
head(q1)
```

```{r}
#query2
head(q2)
```

```{r}
#query6
head(q6)
```


#Exploratory Data Analysis (EDA)
```{r}
#Structure and summary of the loaded data sets
str(uw)        #unexploded
summary(uw)

str(bg)       #background
summary(bg)

#queries
str(q1)
str(q2)
str(q6)
```


```{r}
#unique Brands in unexploded watchlist
unique(uw$Brand)
table(uw$Brand)
```

```{r}
#View Sample Distribution in unexploded watchlist
table(uw$Sample)
#xtabs(~Sample + Brand, uw)
```
#Adding featuring engineered columns into the query data sets
```{r}
# Add derived shape features
add_shape_features <- function(df) {
  df$Circ <- 4 * pi * df$Area / (df$Perim. ^ 2) #Circularity (Circ.)
  df$AR <- df$Major / df$Minor                  #Aspect Ratio
  df$Round <- 1 / df$AR                         #Roundness
  return(df)
}

q1 <- add_shape_features(q1)        #adding to q1
q2 <- add_shape_features(q2)        #adding to q2
q6 <- add_shape_features(q6)        #adding to q6
```


```{r}
#Selecting relevant(7) features and scaling
#Scaling features before clustering as it essential for Mclust
q1_scaled <- scale(q1[, c("Area", "Perim.", "Major", "Minor", "Circ", "AR", "Round")])
q2_scaled <- scale(q2[, c("Area", "Perim.", "Major", "Minor", "Circ", "AR", "Round")])
q6_scaled <- scale(q6[, c("Area", "Perim.", "Major", "Minor", "Circ", "AR", "Round")])
```


#Clusturing using Mclust
```{r}
# Fitting Mclust model
mc_q1 <- Mclust(q1_scaled)
mc_q2 <- Mclust(q2_scaled)
mc_q6 <- Mclust(q6_scaled)

#Summary of models
summary(mc_q1)
summary(mc_q2)
summary(mc_q6)
```


#Principal Component Analysis
#Clusturing charts
```{r}
#PCA1 for q1
pca_q1 <- prcomp(q1_scaled)

fviz_pca_ind(pca_q1, 
             geom.ind = "point", 
             col.ind = as.factor(mc_q1$classification), 
             palette = "jco", 
             addEllipses = TRUE, 
             ellipse.level = 0.95,
             title = "PCA: Query 1 Particles by Mclust Cluster")
```


```{r}
#PCA1 for q2
pca_q2 <- prcomp(q2_scaled)

fviz_pca_ind(pca_q2, 
             geom.ind = "point", 
             col.ind = as.factor(mc_q2$classification), 
             palette = "jco", 
             addEllipses = TRUE, 
             ellipse.level = 0.95,
             title = "PCA: Query 2 Particles by Mclust Cluster")
```


```{r}
#PCA1 for q6
pca_q6 <- prcomp(q6_scaled)

fviz_pca_ind(pca_q6, 
             geom.ind = "point", 
             col.ind = as.factor(mc_q6$classification), 
             palette = "jco", 
             addEllipses = TRUE, 
             ellipse.level = 0.95,
             title = "PCA: Query 6 Particles by Mclust Cluster")
```



```{r}
q1$cluster <- mc_q1$classification
q2$cluster <- mc_q2$classification
q6$cluster <- mc_q6$classification
```


```{r}
# Clean up column names
colnames(uw)[colnames(uw) == "Circ."] <- "Circ"
colnames(bg)[colnames(bg) == "Circ."] <- "Circ"
```


```{r}
#Add a Class Label Column
uw$class <- uw$Brand          # Keep brand name as class
bg$class <- "Background"      # Assign one label to all background particles

```


```{r}
#Select only the seven required features
features <- c("Area", "Perim.", "Major", "Minor", "Circ", "AR", "Round", "class")

uw_clean <- uw[, features]
bg_clean <- bg[, features]
```


```{r}
#Combine into one training data set
train_data <- rbind(uw_clean, bg_clean)

# Convert class column to factor
train_data$class <- as.factor(train_data$class)

# Check class distribution
table(train_data$class)
```


```{r}
# Check for missing values
colSums(is.na(train_data))
```


#Training

```{r}
# Set predictors and outcome
X <- train_data[, c("Area", "Perim.", "Major", "Minor", "Circ", "AR", "Round")]
y <- train_data$class
```


```{r}
#splitting the Data for Training and Validation
set.seed(123)

train_index <- createDataPartition(y, p = 0.8, list = FALSE)
X_train <- X[train_index, ]
X_test <- X[-train_index, ]
y_train <- y[train_index]
y_test <- y[-train_index]
```


```{r}
#Training Random Forest Model
# Get the number of samples per class (balanced down to the smallest class)
table(y_train)
min_class_size <- min(table(y_train))

# Training Random Forest with balanced sampling
set.seed(123)
rf_model <- randomForest(
  x = X_train, 
  y = y_train, 
  ntree = 500,
  importance = TRUE,
  strata = y_train,
  sampsize = rep(min_class_size, length(levels(y_train)))
)

print(rf_model)
```


#Prediction
```{r}
# Predict on test data
rf_preds <- predict(rf_model, newdata = X_test)

# Confusion Matrix
confusionMatrix(rf_preds, y_test)

# Variable Importance Plot
varImpPlot(rf_model)
```



```{r}
query_features <- c("Area", "Perim.", "Major", "Minor", "Circ", "AR", "Round")

# Confirm features exist
stopifnot(all(query_features %in% colnames(q1)))

```


#Prediction for q1,q2,q3
```{r}
# Predict class
q1$predicted_class <- predict(rf_model, q1[, query_features])
q2$predicted_class <- predict(rf_model, q2[, query_features])
q6$predicted_class <- predict(rf_model, q6[, query_features])

# Predict class probabilities (optional confidence)
q1$prediction_confidence <- apply(predict(rf_model, q1[, query_features], type = "prob"), 1, max)
q2$prediction_confidence <- apply(predict(rf_model, q2[, query_features], type = "prob"), 1, max)
q6$prediction_confidence <- apply(predict(rf_model, q6[, query_features], type = "prob"), 1, max)
```


```{r}
head(q1[, c("Area", "Perim.", "Major", "Minor", "predicted_class", "prediction_confidence")])
```


```{r}
#writing the predictions as csv file
write.csv(q1, "q1_with_predictions.csv", row.names = FALSE)
write.csv(q2, "q2_with_predictions.csv", row.names = FALSE)
write.csv(q6, "q6_with_predictions.csv", row.names = FALSE)

```


#Prediction results:
```{r}
# Function to generate a bar plot
plot_bar_distribution <- function(df, query_name) {
  ggplot(df, aes(x = predicted_class)) +
    geom_bar(fill = "#2C3E50") +
    labs(title = paste("Predicted Class Distribution -", query_name),
         x = "Predicted Class", y = "Number of Particles") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Running plots for each query
plot_bar_distribution(q1, "Query 1")
plot_bar_distribution(q2, "Query 2")
plot_bar_distribution(q6, "Query 6")
```


#Piecharts
```{r}
# Function to generate a pie chart
plot_pie_distribution <- function(df, query_name) {
  df_summary <- as.data.frame(table(df$predicted_class))
  colnames(df_summary) <- c("Class", "Count")
  df_summary$Percent <- round(df_summary$Count / sum(df_summary$Count) * 100, 1)
  df_summary$Label <- paste0(df_summary$Class, " (", df_summary$Percent, "%)")

  ggplot(df_summary, aes(x = "", y = Count, fill = Label)) +
    geom_bar(stat = "identity", width = 1) +
    coord_polar("y", start = 0) +
    labs(title = paste("Predicted Class Proportion -", query_name), fill = "Class") +
    theme_void()
}

# Run for each query
plot_pie_distribution(q1, "Query 1")
plot_pie_distribution(q2, "Query 2")
plot_pie_distribution(q6, "Query 6")
```


#Per Clustering Aggregation
#clusterwise prediction
```{r}
# Function to get class distribution per cluster
aggregate_by_cluster <- function(df, query_name) {
  summary <- df %>%
    group_by(cluster, predicted_class) %>%
    summarise(count = n()) %>%
    arrange(cluster, desc(count))

  print(paste("=== Summary for", query_name, "==="))
  print(summary)
  
  return(summary)
}

#summaries
summary_q1 <- aggregate_by_cluster(q1, "Query 1")
summary_q2 <- aggregate_by_cluster(q2, "Query 2")
summary_q6 <- aggregate_by_cluster(q6, "Query 6")
```



```{r}
# Function to plot class distribution per cluster
plot_cluster_composition <- function(summary_df, query_name) {
  ggplot(summary_df, aes(x = factor(cluster), y = count, fill = predicted_class)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(title = paste("Cluster-wise Brand Prediction -", query_name),
         x = "Cluster", y = "Particle Count", fill = "Predicted Class") +
    theme_minimal()
}

plot_cluster_composition(summary_q1, "Query 1")
plot_cluster_composition(summary_q2, "Query 2")
plot_cluster_composition(summary_q6, "Query 6")
```






































































































#Not required
```{r}
# Combine class + one feature for each plot
#ggplot(train_data, aes(x = Area, fill = class)) + 
#  geom_histogram(position = "identity", bins = 50, alpha = 0.4) +
#  labs(title = "Area Distribution by Class")

#ggplot(train_data, aes(x = Circ, fill = class)) + 
#  geom_histogram(position = "identity", bins = 50, alpha = 0.4) +
#  labs(title = "Circularity Distribution by Class")

#ggplot(train_data, aes(x = AR, fill = class)) + 
#  geom_histogram(position = "identity", bins = 50, alpha = 0.4) +
#  labs(title = "Aspect Ratio Distribution by Class")
```

```{r}
#library(MASS)

# Fit LDA
#lda_model <- lda(class ~ Area + Perim. + Major + Minor + Circ + AR + Round, data = train_data)

# View summary
#lda_model
```





